---
title: "Get started with nominatimlite"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started with nominatimlite}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

The goal of `nominatimlite` is to provide a light interface for geocoding
addresses, based on the [Nominatim
API](https://nominatim.org/release-docs/latest/). **Nominatim** is a tool to
search [OpenStreetMap](https://www.openstreetmap.org/) data by name and address
([geocoding](https://wiki.openstreetmap.org/wiki/Geocoding "Geocoding")) and to
generate synthetic addresses of OSM points (reverse geocoding).

This package is derived of the much more mature and complete `tidygeocoder`
package by Jesse Cambon ([link](https://jessecambon.github.io/tidygeocoder/)),
that it is available on CRAN.

## Why `nominatimlite`?

Since `tidygecoder` is much more complete, providing access to several geocoder
services as Mapbox, TomTom or Google, the downloading method is based on `curl`.
In some cases, corporative access to `curl` could be forbidden, so
`nominatimlite` uses an approach to overcome this limitation.

This is the main reason for creating this lite version of `tidygeocoder`.

## What about reverse geocoding?

Reverse geocoding is the process of converting a location as described by
geographic coordinates (latitude, longitude) to a human-readable address or
place name.

This feature is not implemented yet on `nominatimlite`.

## Installation

You can install the developing version of `nominatimlite` with:

``` r
devtools::install_github("dieghernan/nominatimlite")
```

Alternatively, you can install `nominatimlite` using the
[r-universe](https://dieghernan.r-universe.dev/ui#builds):

```r
# Enable this universe
options(repos = c(
    dieghernan = 'https://dieghernan.r-universe.dev',
    CRAN = 'https://cloud.r-project.org'))


install.packages('nominatimlite')
```

## Example

This is a basic example which shows you how to geocode some addresses:

```{r example}
library(nominatimlite)

# Single address
geo_lite("Madrid, Spain")


# Several addresses

geo_lite(c("Madrid", "Barcelona"))

# With options

geo_lite(c("Madrid", "Barcelona"),
  long = "Longitude", lat = "Latitude",
  full_results = TRUE, limit = 5
)



# Restrict search to USA
geo_lite(c("Madrid", "Barcelona"),
  custom_query = list(countrycodes = "US")
)
```

For more advance users, see [Nominatim
docs](https://nominatim.org/release-docs/latest/api/Search/) to check the
parameters available.
