---
title: "Example: nominatimlite and leaflet maps"
subtitle: "Combine nominatimlite and leaflet maps"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  dpi = 300,
  tidy = "styler",
  dev = "ragg_png",
  out.width = "100%"
)
```

## Example

The following example shows how it is possible to create a nice [leaflet
map](https://rstudio.github.io/leaflet/) with data retrieved with
**nominatimlite**.

This widget is browsable and filterable thanks to **crosstalk** and
**reactable**:

```{r example}
# Coffee Shops and Restaurants around the Eiffel Tower


library(nominatimlite)
library(sf)
library(leaflet)
library(dplyr)
library(tidyr)
library(reactable)
library(crosstalk)



# Step 1: Eiffel Tower
eiffel_tower <- geo_lite_sf("Eiffel Tower, Paris, France", points_only = FALSE)



# Step 2: Coffee Shops and Restaurants nearby

# Create a buffer of 1km around the Eiffel Tower
buff <- eiffel_tower %>%
  st_transform(3857) %>%
  st_centroid() %>%
  st_buffer(1000)

cf_bk <- geo_amenity_sf(buff,
  amenity = c("cafe", "restaurant"), limit = 50,
  full_results = TRUE,
  custom_query = list(extratags = TRUE)
) %>%
  # Build address with street, house number, suburb and postcode
  unite("addr", address.road, address.house_number, address.postcode,
    address.suburb,
    sep = ", ", na.rm = TRUE
  )


# Labels and icons
labs <- paste0("<strong>", cf_bk$name, "</strong><br>", cf_bk$addr)

# Assign icons
# Base url for icons
icon_url <- paste0(
  "https://raw.githubusercontent.com/dieghernan/arcgeocoder/",
  "main/vignettes/articles/"
)

leaf_icons <- icons(
  ifelse(cf_bk$type == "cafe",
    paste0(icon_url, "coffee-cup.png"),
    paste0(icon_url, "restaurant.png")
  ),
  iconWidth = 20, iconHeight = 20,
  iconAnchorX = 10, iconAnchorY = 10
)



# Step 3: Crosstalk object
cf_bk_data <- cf_bk %>%
  select(
    Place = name, Type = type, Address = addr,
    City = address.city, URL = extratags.website,
    Phone = extratags.phone
  ) %>%
  SharedData$new(group = "Food")


# Step 4: Leaflet map with crosstalk
# Init leaflet map
lmend <- leaflet(
  data = cf_bk_data,
  elementId = "EiffelTower", width = "100%", height = "60vh",
  options = leafletOptions(minZoom = 12)
) %>%
  addProviderTiles(
    provider = "CartoDB.Positron",
    group = "CartoDB.Positron"
  ) %>%
  addTiles(group = "OSM") %>%
  addPolygons(data = eiffel_tower) %>%
  addMarkers(popup = labs, icon = leaf_icons) %>%
  addLayersControl(
    baseGroups = c("CartoDB.Positron", "OSM"),
    position = "topleft",
    options = layersControlOptions(collapsed = FALSE)
  )



# Step 5: Reactable for filtering
tb <- reactable(cf_bk_data,
  selection = "multiple",
  onClick = "select",
  rowStyle = list(cursor = "pointer"),
  filterable = TRUE,
  searchable = TRUE,
  showPageSizeOptions = TRUE,
  striped = TRUE,
  defaultColDef = colDef(vAlign = "center", minWidth = 150),
  paginationType = "jump",
  elementId = "coffees",
  columns = list(
    Place = colDef(
      sticky = "left", rowHeader = TRUE, name = "",
      cell = function(value) {
        htmltools::strong(value)
      }
    ),
    URL = colDef(cell = function(value) {
      # Render as a link
      if (is.null(value) | is.na(value)) {
        return("")
      }
      htmltools::a(href = value, target = "_blank", as.character(value))
    }),
    Phone = colDef(cell = function(value) {
      # Render as a link
      if (is.null(value) | is.na(value)) {
        return("")
      }
      clearphone <- gsub("-", "", value)
      clearphone <- gsub(" ", "", clearphone)
      htmltools::a(
        href = paste0("tel:", clearphone), target = "_blank",
        as.character(value)
      )
    })
  )
)
```

## Widget

```{r widget}
# Last step: Display all
htmltools::browsable(
  htmltools::tagList(lmend, tb)
)
```

## Attributions

-   [Eiffel tower icons created by Freepik -
    Flaticon](https://www.flaticon.com/free-icons/eiffel-tower "eiffel tower icons")
-   [Mug icons created by Freepik -
    Flaticon](https://www.flaticon.com/free-icons/mug "mug icons")
-   [Food icons created by Freepik -
    Flaticon](https://www.flaticon.com/free-icons/food "restaurant icons")

## Session info

<details>

<summary>Details</summary>

```{r session, echo=FALSE}
if (!require("sessioninfo")) {
  install.packages("sessioninfo")
}
sessioninfo::session_info()
```

</details>
